---
title: "p8105_hw1_mx2286.Rmd"
author: "Mingye"
date: "2024-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



####Question 1
```{r}
# Load necessary packages
library(palmerpenguins)
data("penguins", package = "palmerpenguins")
str(penguins)

summary(penguins)

```
**Disscussion**
Penguins Dataset Overview

The `penguins` dataset contains information about penguin species collected from Palmer Station, Antarctica. Important variables in this dataset include:

- **Species**: Species of penguin (`Adelie`, `Chinstrap`, `Gentoo`)
- **island**: Island name (`Biscoe`, `Dream`, `Torgersen`)
- **bill_length_mm**: Length of the penguin's bill in millimeters
- **bill_depth_mm**: Depth of the penguin's bill in millimeters
- **flipper_length_mm**: Length of the penguin's flippers in millimeters
- **body_mass_g**: Body mass in grams
- **sex**: Sex of the penguin (`male`, `female`)
- **year**: Time when the data recorded

The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length across all penguins is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

**Conclusion**
This dataset provides valuable biological and ecological data for analysis, enabling the study of penguin species and their physical characteristics in relation to their environment. 


```{r}
library(ggplot2)

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Flipper Length vs. Bill Length by Penguin Species",
       x = "Bill Length (mm)", y = "Flipper Length (mm)") +
  theme_minimal()

ggsave("penguin_scatterplot.png")
```


####Question 2

---
title: "Data Frame and Variable Type Coercion"
output: html_document
---

```{r}
set.seed(123) # For reproducibility

# Create a random sample from a standard normal distribution
random_sample <- rnorm(10)

# Create a logical vector indicating whether the elements of random_sample are greater than 0
logical_vector <- random_sample > 0

# Create a character vector representing colors
character_vector <- c("Red", "Yellow", "Red", "Brown", "Purple", 
                      "Purple", "Green", "Green", "Yellow", "Green")

# Create a factor vector with 3 levels
factor_vector <- factor(rep(c("A", "B", "C"), length.out = 10))

# Combine the vectors into a data frame
df <- data.frame(random_sample, logical_vector, character_vector, factor_vector)

# Print the data frame
print(df)

# Load the tidyverse package
library(tidyverse)

# Calculate the mean of the numeric vector (random_sample)
mean_random_sample <- mean(pull(df, random_sample))
mean_random_sample  # This works because random_sample is numeric

# Calculate the mean of the logical vector (logical_vector)
mean_logical_vector <- mean(pull(df, logical_vector))
mean_logical_vector  # This works because TRUE is treated as 1 and FALSE as 0

# Attempt to calculate the mean of the character vector (character_vector)
# This will fail because character data cannot be averaged
# Commenting out to avoid error
# mean_character_vector <- mean(pull(df, character_vector))

# Attempt to calculate the mean of the factor vector (factor_vector)
# This will also fail because factors are categorical, not numeric
# Commenting out to avoid error
# mean_factor_vector <- mean(pull(df, factor_vector))

# Convert logical vector to numeric
numeric_logical_vector <- as.numeric(df$logical_vector)
numeric_logical_vector  # Output shows that TRUE becomes 1 and FALSE becomes 0.

# Convert character vector to numeric
numeric_character_vector <- as.numeric(as.factor(df$character_vector))
numeric_character_vector  # Characters are converted to their underlying factor levels.

# Convert factor vector to numeric
numeric_factor_vector <- as.numeric(df$factor_vector)
numeric_factor_vector  # Factor levels are converted to their underlying numeric codes.
```
### Explanation of What Happens:

**Logical Vector Conversion**:
   Logical vectors in R can be easily coerced to numeric, with `TRUE` being converted to 1 and `FALSE` to 0. This is because logical values are essentially treated as binary numbers in R.

**Character Vector Conversion**:
   When you created the data frame, R likely converted the character vector to a factor. As a result, `as.numeric()` on this vector returns the internal numeric codes of the factor levels (1 for the first unique character, 2 for the second, and so on). This explains why it outputs numeric values even though the original data is not numeric.

**Factor Vector Conversion**:
   Factors are stored as integers internally, where each level corresponds to a numeric code (e.g., 1 for the first level, 2 for the second level, etc.). When you apply `as.numeric()` to a factor, it returns these underlying codes rather than the factor labels themselves.

### Conclusion:

**Logical vector**: Converts logically and correctly to 1s and 0s.
**Character vector**: Was treated as a factor internally, leading to numeric output corresponding to factor levels.
**Factor vector**: As expected, returns the numeric codes for each factor level.
